### 2014-08-08 Studies/2012_Chicago/Report/Chicago_Fire.tex
Chicago: minor update to report

On November 2, 2012, a fire in a 2~\sfrac{1}{2}~story wood frame residential structure claimed the life of a captain of the Chicago Fire Department. NIST examined the fire dynamics of this incident at the request of NIOSH and the Chicago Fire Department. Computer simulations of the fire incident were conducted using the NIST Fire Dynamics Simulator~\cite{FDS_Users_Guide} and Smokeview~\cite{Smokeview_Users_Guide}  software to provide insight into the fire development and thermal conditions that likely existed in the residence during the fire. The specific objectives of the simulations detailed in this report are: 
\begin{enumerate}
\item To examine the effect of the flow path (including temperature, pressure, and fire conditions) in this multi-story residential structure using physics-based calculations.
\item To provide visualizations of the fire behavior that are representative of the conditions that members of the Chicago Fire Department likely experienced during the course of their interior operations.
\end{enumerate}
This document describes the input and the results of the FDS (version 6.0.0) simulations. The simulations were developed using a combination of knowledge of the fire scenario and appropriate engineering approximations and assumptions. Analysis of the simulation results focuses on the hazardous conditions that developed following the development of a flow path inside of the structure. This document is organized as follows: Section~\ref{fire_sum} provides a summary of the fire incident, Section~\ref{model} describes the relevant model input parameters and assumptions that were used to develop the simulations, Section~\ref{results} presents the simulation results, and Section~\ref{discussion} discusses the simulation results as they relate to firefighter safety and effectiveness.

@@

\chapter{Summary}
NIST's Fire Dynamics Simulator was used to provide insight into the fire dynamics within a 2~\sfrac{1}{2}~story single family structure in Chicago, IL, that resulted in the death the captain of a fire department. The fuel, fire size, and fire growth rate that were used in the FDS simulations were estimated by taking into account all of the available information including the NIOSH post-incident report~\cite{NIOSH:Bowyer}, post-incident pictures, videos, and relevant literature. This resulted in a specified source fire of 12.3~MW. However, based on the limited ventilation conditions in the attic and enclosed porch, the FDS simulation results indicated a maximum HRR of approximately 9~MW as a result of the limited supply of oxygen.

The fire started in the attic and spread down into the enclosed porch on the rear of the structure. The spread of fire and hot gases into the second floor of the structure was limited by a closed steel-faced, wood-framed door that was located between the enclosed porch and the hallway. After exposure to the elevated temperatures and pressure from the fire in the porch area, the wood frame of the door decomposed and the steel faces of the door failed and collapsed inward. The door failure resulted in the establishment of a flow path between the higher pressure and higher temperature conditions in the enclosed porch and the lower pressure and lower temperature conditions in the hallway and kitchen areas of the second floor. The temperature of the gases that flowed into the hallway at face height exceeded 260~$^{\circ}$C (500~$^{\circ}$F). Unfortunately, a captain and firefighter were advancing a hose line in the hallway at the time of the door failure. The captain was overcome by the rapid change in conditions before he could exit to the safety of the kitchen area. After a Mayday call, the captain was rescued from the structure but later succumbed to his injuries.

@@

\chapter{Dimensioned Drawings}

\begin{figure}[!ht]
\centering
\includegraphics[width=1\textwidth]{../Figures/50th_Place_1st_Floor}
\caption[Dimensioned drawing of first floor.]{Dimensioned drawing of first floor. The measurements are accurate to within 15~cm (6~in).}
\label{fig:first_floor}
\end{figure}

>>>>

On November 2, 2012, a fire in a 2~\sfrac{1}{2}~story wood frame residential structure claimed the life of a captain of the Chicago Fire Department. NIST examined the fire dynamics of this incident at the request of NIOSH and the Chicago Fire Department. Computer simulations of the fire incident were conducted using the NIST Fire Dynamics Simulator~\cite{FDS_Users_Guide} and Smokeview~\cite{Smokeview_Users_Guide}  software to provide insight into the fire development and thermal conditions that likely existed in the residence during the fire. The specific objectives of the simulations detailed in this report are: 
\begin{enumerate}
\item To examine the effect of the flow path (including temperature, pressure, and fire conditions) in this multi-story residential structure using physics-based calculations.
\item To provide visualizations of the fire behavior that are representative of the conditions that members of the Chicago Fire Department likely experienced during the course of their interior operations.
\end{enumerate}
This document describes the input and the results of the FDS (version 6.0.0) simulations. The simulations were developed using a combination of knowledge of the fire scenario and appropriate engineering approximations and assumptions. Analysis of the simulation results focuses on the hazardous conditions that developed following the development of a flow path inside of the structure. This document is organized as follows: Section~\ref{fire_sum} provides a summary of the fire incident, Section~\ref{model} describes the relevant model input parameters and assumptions that were used to develop the simulations, Section~\ref{results} presents the simulation results, and Section~\ref{discussion} discusses the simulation results as they relate to firefighter safety and effectiveness. Appendinx~\ref{app_draw} contains dimensioned drawings of the first and second floor of the structure.

@@

\chapter{Summary}
NIST's Fire Dynamics Simulator was used to provide insight into the fire dynamics within a 2~\sfrac{1}{2}~story single family structure in Chicago, IL, that resulted in the death the captain of a fire department. The fuel, fire size, and fire growth rate that were used in the FDS simulations were estimated by taking into account all of the available information including the NIOSH post-incident report~\cite{NIOSH:Bowyer}, post-incident pictures, videos, and relevant literature. This resulted in a specified source fire of 12.3~MW in the attic and rear porch of the structure. However, based on the limited ventilation conditions in the attic and enclosed porch, the FDS simulation results indicated a maximum HRR of approximately 9~MW as a result of the limited supply of oxygen.

The fire started in the attic and spread down into the enclosed porch on the rear of the structure. The spread of fire and hot gases into the second floor of the structure was limited by a closed steel-faced, wood-framed door that was located between the enclosed porch and the hallway. After exposure to the elevated temperatures and pressure from the fire in the porch area, the wood frame of the door decomposed and the steel faces of the door failed and collapsed inward. The door failure resulted in the establishment of a flow path between the higher pressure and higher temperature conditions in the enclosed porch and the lower pressure and lower temperature conditions in the hallway and kitchen areas of the second floor. The temperature of the gases that flowed into the hallway at face height was calculated to exceed 260~$^{\circ}$C (500~$^{\circ}$F). Unfortunately, a captain and firefighter were advancing a hose line in the hallway at the time of the door failure. The captain was overcome by the rapid change in conditions before he could exit to the safety of the kitchen area. After a Mayday call, the captain was rescued from the structure but later succumbed to his injuries.

@@

\chapter{Dimensioned Drawings}
\label{app_draw}
\begin{figure}[!ht]
\centering
\includegraphics[width=1\textwidth]{../Figures/50th_Place_1st_Floor}
\caption[Dimensioned drawing of first floor.]{Dimensioned drawing of first floor. The measurements are accurate to within 15~cm (6~in).}
\label{fig:first_floor}
\end{figure}
